---
title: "New STR mutations in BXD"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
      df_print: paged
      code_folding: hide
      toc: true
      toc_float: true
      number_sections: true
---

<!-- Set with output width -->

```{=html}
<style type="text/css">
.main-container {
    max-width: 1600px;
    margin-left: auto;
    margin-right: auto;
}
h1 {font-size: 2em}
h2 {font-size: 1.5em}
h3 {font-size: 1.17em}
h4 {font-size: 1.12em}
h5 {font-size: 0.83em}
h6 {font-size: 0.75em}
</style>
```
```{r setup, include=FALSE}
	# options
	knitr::opts_chunk$set(echo = TRUE)
	options(stringsAsFactors = FALSE)
	# libraries
	library(fs)
	library(tidyverse)
  library(devtools)
	library(cowplot)
	library(GGally)
	library(ggbeeswarm)
  library(progress)
	library(qtl2)
	library(readr)
	library(ggrepel)

	# Global configs
	data_dir = '../data/'
  n_permute = 100
	#cache_dir = '../data/analysis_cache/bxd_qtl_scans'; dir_create(cache_dir)
	
	denovo_strs = readr::read_csv('../outs/denovo_strs_filtered.csv')
	strain_info = readr::read_csv('../outs/strain_info.csv')
  gtloc_per_strain = readr::read_csv('../outs/gtloc_per_strain.csv')
  motif_info = readr::read_csv('../outs/motif_info.csv')
  qtl_data = readRDS('../outs/qtl_data.rds')
  # load function for calculating phenotypes
	source('../utils/calc_pheno_vals.R')

  # define color scheme for motifs
	motif_colors = c(
		'all' = '#000000', 
		'1'   = '#989896', 
		'2'   = '#FF2900', 
		'3'   = '#E7D707', 
		'4'   = '#3F369B', 
		'5'   = '#006D4A', 
		'6'   = '#D12C9C')
  
	phenos = c(
		'% denovo'           = 'denovo_perc_abundance',
		'delta (RU) expan'   = 'expand_delta_ru',
		'delta (RU) contr'   = 'contract_delta_ru',
		'% expanded'         = 'proportion_expanded'
	)
	
	FINALML = 'all' # final motif length. either "4" (tetranucs only) or "all" (all motifs)
```


# Load data 

#. Attach motif info to new variant str information
#. Bundle epoch1b and epoch1c together with epoch1a
#. Fix BXD029 gen_inbreeding

```{r fig.width = 6, fig.height = 4, fig.align = 'center', cache = TRUE, eval = TRUE}
	# join motif length to new variant STRs (note: these are homozygous)
	denovo_strs = denovo_strs %>%
	left_join(motif_info %>% select(chr, pos, end, motif_len), by = c('chr', 'pos', 'end'))
	strain_info = strain_info %>%
	  mutate(off_epoch = recode(off_epoch, epoch_1a = 'epoch_1', epoch_1b = 'epoch_1', epoch_1c = 'epoch_1')) 
	num_mut = nrow(denovo_strs %>% distinct(chr, pos, end))
  print(paste("Num mutations:", num_mut))
  print(paste("Mutation length:", FINALML))
```

# Calculate all phenotypes for different filtering params
```{r fig.width = 6, fig.height = 4, fig.align = 'center', cache = TRUE, eval = TRUE}
  # calculate %expanded for different motif lengths and filtering parameters
	mls = c('all', 2, 3, 4, 5, 6) %>% set_names(.) # motif len
	mpps = c(0, 5, 10, 20, 50) %>% set_names(.) # min number of total mutations per strain to include
	mdsls = c(1000, 50, 10, 3, 1) %>% set_names(.) # max_denovo_strains_per_loc
	
	all_phenos = 
	  map_df(phenos, function(pheno) {
	    map_df(mls, function(ml) {
			  map_df(mpps, function(min_pts_per_phe) {
				  map_df(mdsls, function(max_denovo_strains_per_loc) {
					  calc_pheno_vals(denovo_strs, pheno, 
									min_pts_per_phe = min_pts_per_phe, 
									max_denovo_strains_per_loc = max_denovo_strains_per_loc, 
									ml = ml,
									gtloc_per_strain = gtloc_per_strain %>% dplyr::rename(n_gt = n_loci)) %>%
						mutate(motif_len = as.character(ml))
				}, .id = 'max_denovo_strains_per_loc')
			}, .id = 'min_pts_per_phe')
		}, .id='motif_len')
	}, .id='metric')
	
	pheno_vals = all_phenos %>% 
	  filter(min_pts_per_phe == 10, max_denovo_strains_per_loc == 10, motif_len == '4') %>%
	  select(strain, motif_len, pheno, n, metric)
```


# GW scan all phenotypes with loc/strain filtering

#. min_pts_per_phe == 10, max_denovo_strains_per_loc == 10

```{r fig.width = 16, fig.height = 8, fig.align = 'center', cache = TRUE, eval = TRUE}
		# run QTL mapping using SNP founder states and final calculated phenotype values
		# all chromosomes
		a_thresh = 0.05
		# set strains
		strains = strain_info %>% filter(off_epoch != 'founder') %>% pull(bxd_id) %>% unique
		
		final_phenos_qtl = map(names(phenos), function(pt) {
		  print(pt)
			# subset phenotype values
		  this_phenos = all_phenos %>%
		    filter(min_pts_per_phe == 10, max_denovo_strains_per_loc == 10, motif_len == FINALML) %>%
		    filter(metric == pt) %>%
		    filter(strain %in% strains) %>%
		  	select(strain, pheno) %>% 
				column_to_rownames(var = 'strain') %>% as.matrix
			# check if enough strains
			pheno_strains = row.names(this_phenos)
			if (length(pheno_strains) < 5) return(NULL)
			
			this_probs = qtl_data$snp_probs[pheno_strains,]
			this_kinship = qtl_data$snp_kinship
			
			this_covar = strain_info %>% 
				  select(bxd_id, gen_inbreeding) %>%
				  filter(bxd_id %in% pheno_strains) %>%
				column_to_rownames(var = 'bxd_id') %>% as.matrix
			
			# run qtl analysis
			qtl_res = scan1(genoprobs = this_probs, 
							pheno = this_phenos,
							kinship = this_kinship,
							addcovar = this_covar
			)
			perm_res = scan1perm(genoprobs = this_probs, 
								 pheno = this_phenos,
								 kinship = this_kinship,
								 addcovar = this_covar,
								 n_perm = n_permute
			)
			# output
			return(list(qtl_res = qtl_res, perm_res = perm_res))
		})
		names(final_phenos_qtl) = names(phenos)
		# make tidy
		final_qtl_res = map_df(final_phenos_qtl, function(.x) {
			as.data.frame(.x$qtl_res) %>%
				rownames_to_column(var = 'marker') %>%
				as_tibble %>%
				dplyr::rename(LOD = pheno) %>%
				mutate(lod_thresh = quantile(.x$perm_res[,'pheno'], probs = 1-a_thresh, names = FALSE)) %>%
				# left_join(snp_phys_map %>% select(marker, chr, pos, end), by = 'marker')
				separate('marker', c('chr', 'pos', 'end'), convert = TRUE)
		}, .id = 'metric')
		write.csv('../outs/final_qtl_res.csv')
		# Plot QTL signal for all phenotypes
		p = final_qtl_res %>%
		arrange(match(metric, names(phenos))) %>%
		mutate(delta_dir = if_else(str_detect(metric, ' (expan|contr)$'),
								   str_replace(metric, '.* (expan|contr)$', '\\1'), 
								   'all')) %>%
		mutate(metric = if_else(delta_dir != 'all', 
								str_replace(metric, ' (expan|contr)$', ''),
								metric)) %>%
		mutate(metric = fct_inorder(metric)) %>%
		# recode labels for final figure
		mutate(metric = recode(metric, !!!c('% denovo' = 'mutation count', 
											'delta (RU)' = 'expansion size', 
											'% expanded' = 'expansion propensity'))) %>%
		mutate(chr = str_replace(chr, 'chr', '')) %>%
		mutate(chr = fct_relevel(chr, str_sort(unique(chr), numeric = TRUE))) %>%
		# filter(metric %in% c('% denovo', '% expanded', 'delta (RU) expan')) %>%
		mutate(across(c(pos, end), ~.x/1e6)) %>%
		ggplot(aes(pos, LOD, color = delta_dir)) + 
		geom_step() +
		geom_hline(data = ~.x %>% distinct(metric, chr, lod_thresh, delta_dir),
			   aes(yintercept = lod_thresh, color = delta_dir), linetype = 'dashed') + 
		facet_grid(metric~chr, scales = 'free_x', switch = 'x') + 
		scale_x_continuous(breaks = scales::breaks_pretty(n = 2), 
						   guide = guide_axis(angle = 60)) +
		scale_color_brewer(palette = 'Dark2', guide = guide_legend(title = NULL)) + 
		# coord_cartesian(ylim = c(0, 4)) +
		theme_half_open() +
		theme(
			panel.spacing.x = unit(0, 'pt'),
			axis.title.x = element_blank(),
			strip.placement = 'outside',
			strip.text.y = element_text(angle = 90),
			panel.border = element_rect(color = 'gray70'),
			axis.text.x = element_blank(),
			axis.ticks.x = element_blank(),
			legend.position = 'top'
		)
		p
		ggsave(fs::path('pdfs', 'Fig2_GWScan.pdf'), p, w = 6, h = 4)
```

# GW scan all phenotypes with loc/strain filtering - TETRA

#. min_pts_per_phe == 10, max_denovo_strains_per_loc == 10

```{r fig.width = 16, fig.height = 8, fig.align = 'center', cache = TRUE, eval = TRUE}
		# run QTL mapping using SNP founder states and final calculated phenotype values
		# all chromosomes
		a_thresh = 0.05
		# set strains
		strains = strain_info %>% filter(off_epoch != 'founder') %>% pull(bxd_id) %>% unique
		
		final_phenos_qtl_tetra = map(names(phenos), function(pt) {
		  print(pt)
			# subset phenotype values
		  this_phenos = all_phenos %>%
		    filter(min_pts_per_phe == 10, max_denovo_strains_per_loc == 10, motif_len == '4') %>%
		    filter(metric == pt) %>%
		    filter(strain %in% strains) %>%
		  	select(strain, pheno) %>% 
				column_to_rownames(var = 'strain') %>% as.matrix
			# check if enough strains
			pheno_strains = row.names(this_phenos)
			if (length(pheno_strains) < 5) return(NULL)
			
			this_probs = qtl_data$snp_probs[pheno_strains,]
			this_kinship = qtl_data$snp_kinship
			
			this_covar = strain_info %>% 
				  select(bxd_id, gen_inbreeding) %>%
				  filter(bxd_id %in% pheno_strains) %>%
				column_to_rownames(var = 'bxd_id') %>% as.matrix
			
			# run qtl analysis
			qtl_res = scan1(genoprobs = this_probs, 
							pheno = this_phenos,
							kinship = this_kinship,
							addcovar = this_covar
			)
			perm_res = scan1perm(genoprobs = this_probs, 
								 pheno = this_phenos,
								 kinship = this_kinship,
								 addcovar = this_covar,
								 n_perm = n_permute
			)
			# output
			return(list(qtl_res = qtl_res, perm_res = perm_res))
		})
		names(final_phenos_qtl_tetra) = names(phenos)
		# make tidy
		final_qtl_res_tetra = map_df(final_phenos_qtl_tetra, function(.x) {
			as.data.frame(.x$qtl_res) %>%
				rownames_to_column(var = 'marker') %>%
				as_tibble %>%
				dplyr::rename(LOD = pheno) %>%
				mutate(lod_thresh = quantile(.x$perm_res[,'pheno'], probs = 1-a_thresh, names = FALSE)) %>%
				# left_join(snp_phys_map %>% select(marker, chr, pos, end), by = 'marker')
				separate('marker', c('chr', 'pos', 'end'), convert = TRUE)
		}, .id = 'metric')
		write.csv('../outs/final_qtl_res_tetra.csv')

		# Plot QTL signal for all phenotypes
		p = final_qtl_res_tetra %>%
		arrange(match(metric, names(phenos))) %>%
		mutate(delta_dir = if_else(str_detect(metric, ' (expan|contr)$'),
								   str_replace(metric, '.* (expan|contr)$', '\\1'), 
								   'all')) %>%
		mutate(metric = if_else(delta_dir != 'all', 
								str_replace(metric, ' (expan|contr)$', ''),
								metric)) %>%
		mutate(metric = fct_inorder(metric)) %>%
		# recode labels for final figure
		mutate(metric = recode(metric, !!!c('% denovo' = 'mutation count', 
											'delta (RU)' = 'expansion size', 
											'% expanded' = 'expansion propensity'))) %>%
		mutate(chr = str_replace(chr, 'chr', '')) %>%
		mutate(chr = fct_relevel(chr, str_sort(unique(chr), numeric = TRUE))) %>%
		# filter(metric %in% c('% denovo', '% expanded', 'delta (RU) expan')) %>%
		mutate(across(c(pos, end), ~.x/1e6)) %>%
		ggplot(aes(pos, LOD, color = delta_dir)) + 
		geom_step() +
		geom_hline(data = ~.x %>% distinct(metric, chr, lod_thresh, delta_dir),
			   aes(yintercept = lod_thresh, color = delta_dir), linetype = 'dashed') + 
		facet_grid(metric~chr, scales = 'free_x', switch = 'x') + 
		scale_x_continuous(breaks = scales::breaks_pretty(n = 2), 
						   guide = guide_axis(angle = 60)) +
		scale_color_brewer(palette = 'Dark2', guide = guide_legend(title = NULL)) + 
		# coord_cartesian(ylim = c(0, 4)) +
		theme_half_open() +
		theme(
			panel.spacing.x = unit(0, 'pt'),
			axis.title.x = element_blank(),
			strip.placement = 'outside',
			strip.text.y = element_text(angle = 90),
			panel.border = element_rect(color = 'gray70'),
			axis.text.x = element_blank(),
			axis.ticks.x = element_blank(),
			legend.position = 'top'
		)
		p
		ggsave(fs::path('pdfs', 'Fig2_GWScan-Tetra.pdf'), p, w = 6, h = 4)
```
# Visualize scan sensitivity results %expanded

## Min-points per phenotype and max strains per denovo filtering

#. All strains used
#. Robust to filtering conditions and using STRs with different motif lengths
#. Difference by `max_denovo_strains_per_loc`
#. Small difference by `min_pts_per_phe`

```{r echo=TRUE, fig.width = 6, fig.height = 4, fig.align = 'center', cache = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
chroms = c('13')
# set strains
strains = strain_info %>% pull(bxd_id) %>% unique
a_thresh = 0.05
perc_expand_final_qtl = all_phenos %>%
  filter(metric == '% expanded') %>%
  nest(pheno_vals = c(strain, pheno, n)) %>%
  group_by(min_pts_per_phe, max_denovo_strains_per_loc, motif_len, metric) %>%
  summarise(qtl_res = map(pheno_vals, function(.x) {
					.x = .x %>% filter(strain %in% strains)
				# transform the 
				this_phenos = .x %>%
					select(strain, pheno) %>% 
					column_to_rownames(var = 'strain') %>% as.matrix
				# get strains for which we actually have a phenotype
				pheno_strains = .x %>% pull(strain)
				if (length(pheno_strains) < 5) return(NULL)
				# subset snp probabilities and kinship matrices
				this_probs = qtl_data$snp_probs[pheno_strains,chroms]
				this_kinship = qtl_data$snp_kinship[chroms]
				this_covar = strain_info %>% 
					select(bxd_id, gen_inbreeding) %>% 
					filter(bxd_id %in% pheno_strains) %>%
					column_to_rownames(var = 'bxd_id') %>% as.matrix
				# run qtl analysis
				qtl_res = scan1(genoprobs = this_probs, 
								pheno = this_phenos,
								kinship = this_kinship,
								addcovar = this_covar
				)
				perm_res = scan1perm(genoprobs = this_probs, 
									 pheno = this_phenos,
									 kinship = this_kinship,
									 addcovar = this_covar,
									 n_perm = n_permute
				)
				# combine
				qtl_res = as.data.frame(qtl_res) %>%
					rownames_to_column(var = 'marker') %>%
					as_tibble %>%
					rename(LOD = pheno) %>%
					mutate(lod_thresh = quantile(perm_res[,'pheno'], probs = 1-a_thresh, names = FALSE)) %>%
					# left_join(snp_phys_map %>% select(marker, chr, pos, end), by = 'marker')
					separate('marker', c('chr', 'pos', 'end'), convert = TRUE)
				# output
				return(qtl_res)
			}), .groups = 'drop')
		perc_expand_final_qtl = perc_expand_final_qtl %>% unnest(cols = qtl_res)
		
		###### visualize
		# loop over chromosomes
	  by_chrom = map2(list('chr13') %>% set_names(.), 
					list(c(0, 10), c(0, 4)), 
		  function(chrom, ylims) {
	    # made a list of plots
	    p_list = perc_expand_final_qtl %>% 
		  filter(chr == chrom) %>%
		  mutate(
			  motif_len = fct_relevel(motif_len, mls),
			  min_pts_per_phe = fct_relevel(min_pts_per_phe, as.character(mpps)) %>% fct_relabel(~str_c('n>=', .x)),
			  max_denovo_strains_per_loc = fct_relevel(max_denovo_strains_per_loc, as.character(mdsls)) %>% fct_relabel(~str_c('n<=', .x))
		) %>%
		  nest(data = c(motif_len, chr, pos, end, LOD, lod_thresh)) %>%
		  arrange(min_pts_per_phe, max_denovo_strains_per_loc) %>%
		  mutate(grid_pos = 1:n()) %>%
		  nest(cond = c(grid_pos, min_pts_per_phe, max_denovo_strains_per_loc)) %>%
		  mutate(r_id = 1:n()) %>%
		  group_by(r_id) %>%
		  mutate(p = map2(cond, data, function(cond, data) {
		    p = data %>%
				mutate(across(c(pos, end), ~.x/1e6)) %>%
				ggplot(aes(pos, LOD, color = motif_len)) +
				geom_step() +
				geom_text_repel(data = ~.x %>% 
							group_by(motif_len) %>% 
							slice_max(n = 1, order_by = LOD),
						aes(label = sprintf('%0.1f', LOD), x = pos, y = LOD), color = 'black') +
				geom_hline(data = ~.x %>% distinct(motif_len, lod_thresh),
					   aes(yintercept = lod_thresh), linetype = 'dashed') + 
				facet_wrap(~motif_len, nrow = 1, strip.position = 'bottom', drop = FALSE) +
				scale_y_continuous(expand = expansion(0.01, 0)) + 
				scale_color_brewer(palette = 'Paired', drop = FALSE) + 
				coord_cartesian(ylim = ylims) +
				labs(y = cond %>% pull(min_pts_per_phe), 
				 title = cond %>% pull(max_denovo_strains_per_loc)) +
				theme_half_open() +
				theme(axis.text.x = element_blank(),
				  axis.ticks.x = element_blank(),
				  axis.title.x = element_blank(),
				  strip.background = element_blank(),
				  strip.placement = 'outside',
				  axis.title.y = element_text(size = 12, face = 'bold'),
				  plot.title = element_text(hjust = 0.5, size = 12),
				  panel.spacing = unit(2, 'pt'),
				  plot.margin = margin(0.1, 0.1, 0.1, 0.1, 'cm'),
				  legend.position = 'none')
			if (cond$grid_pos > 5) p = p + theme(plot.title = element_blank())
			if (cond$grid_pos %% 5 != 1) p = p + theme(axis.title.y = element_blank(), 
								 axis.text.y = element_blank(), 
								 axis.ticks.y = element_blank(),
								 axis.line.y = element_blank())
			if (cond$grid_pos < 21) p = p + theme(strip.text = element_blank())
			p
		}))
	p = plot_grid(plotlist = p_list$p, nrow = 5, ncol = 5, 
				  rel_heights = c(1.1, 0.9, 0.9, 0.9, 1.1),
				  rel_widths = c(1.1, 0.9, 0.9, 0.9, 0.9)
	)
	return(p)
	})
	p = by_chrom[[1]]
	p
	ggsave(fs::path('pdfs', 'Suppl_FigSensitivityScan.pdf'), p, w = 16, h = 16)
```

# Visualize %expanded by motif length (no filtering)

```{r fig.width = 6, fig.height = 4, fig.align = 'center', cache = TRUE, eval = TRUE}
	# visualize perc expanded. Note this is with *no filtering*
	p = all_phenos %>%
		filter(min_pts_per_phe == 0, max_denovo_strains_per_loc == 1000) %>%
		filter(metric == '% expanded') %>%
		mutate(motif_len = fct_relevel(motif_len, c('all', 2:6))) %>%
		ggplot(aes(motif_len, pheno, color = motif_len)) +
		geom_quasirandom(dodge.width = 0.9) +
		geom_boxplot(aes(group = motif_len), 
					 fill = 'white', color = 'black', 
					 width = 0.2, outlier.shape = NA,
					 position = position_dodge(width = 0.9)) + 
		# scale_color_brewer(palette = 'Dark2', guide = guide_legend(title = NULL)) + 
		scale_color_manual(
			values = motif_colors,
			guide = guide_legend(title = NULL, nrow = 1)
		) + 
		theme_half_open() +
		theme(
			strip.text.y = element_text(angle = 0),
			# axis.text.y = element_blank(),
			# axis.ticks.y = element_blank(),
			# axis.title.y = element_blank(),
			legend.position = 'top'
		)
	p
	ggsave(fs::path('pdfs', 'Suppl_FigPercExpandedDistribution.pdf'), p, w = 6, h = 4)
```

# Visualize QTL signal by epoch

epoch3b has the most power to resolve association, but likely present in all strains

```{r echo=TRUE, fig.width = 16, fig.height = 8, fig.align = 'center', cache = TRUE, eval = TRUE}
# run QTL mapping separately for different epochs
epoch_grps = strain_info %>%
		select(strain = bxd_id, off_epoch) %>%
		mutate(off_epoch = str_replace(off_epoch, 'epoch_', '')) %>%
		filter(off_epoch != 'founder') %>%
		mutate(epoch_grp = if_else(off_epoch %in% 4:7, '4-7', off_epoch))
# add the 2+3a group
epoch_grps = epoch_grps %>%
		bind_rows(epoch_grps %>% filter(epoch_grp %in% c('2', '3b')) %>% mutate(epoch_grp = '2,3b')) %>%
		bind_rows(epoch_grps %>% mutate(epoch_grp = 'all'))
	
perc_expand_by_epoch_final = epoch_grps %>%
			split(.$epoch_grp) %>%
			map('strain') %>%
			map_df(function(strains) {
				# transform the 
				this_phenos = all_phenos %>%
				  filter(metric == '% expanded') %>%
					filter(min_pts_per_phe == 10, 
						   max_denovo_strains_per_loc == 10,
						   motif_len == FINALML) %>%
					filter(strain %in% strains) %>%
					select(strain, pheno) %>% 
					column_to_rownames(var = 'strain') %>% as.matrix
				# get strains for which we actually have a phenotype
				pheno_strains = row.names(this_phenos)
				if (length(pheno_strains) < 5) return(NULL)
				# subset snp probabilities and kinship matrices
				this_probs = qtl_data$snp_probs[pheno_strains,]
				this_kinship = qtl_data$snp_kinship[]
				this_covar = strain_info %>% 
					select(bxd_id, gen_inbreeding) %>% 
					filter(bxd_id %in% pheno_strains) %>%
					column_to_rownames(var = 'bxd_id') %>% as.matrix
				# run qtl analysis
				qtl_res = scan1(genoprobs = this_probs, 
								pheno = this_phenos,
								kinship = this_kinship,
								addcovar = this_covar
				)
				perm_res = scan1perm(genoprobs = this_probs, 
									 pheno = this_phenos,
									 kinship = this_kinship,
									 addcovar = this_covar,
									 n_perm = n_permute
				)
				# combine
				qtl_res = as.data.frame(qtl_res) %>%
					rownames_to_column(var = 'marker') %>%
					as_tibble %>%
					rename(LOD = pheno) %>%
					mutate(lod_thresh = quantile(perm_res[,'pheno'], probs = 1-a_thresh, names = FALSE)) %>%
					# left_join(snp_phys_map %>% select(marker, chr, pos, end), by = 'marker')
					separate('marker', c('chr', 'pos', 'end'), convert = TRUE)
				# output
				return(qtl_res)
			}, .id = 'epoch_grp')
  write.csv('../outs/perc_expand_by_epoch_final.csv')
	p = perc_expand_by_epoch_final %>%
		mutate(epoch_grp = fct_relevel(epoch_grp, c('all', '1', '2', '3a', '3b', '4-7', '2,3b'))) %>%
		filter(epoch_grp != '2,3b') %>%
		mutate(chr = fct_relevel(chr, str_sort(unique(chr), numeric = TRUE))) %>%
		mutate(across(c(pos, end), ~.x/1e6)) %>%
		ggplot(aes(pos, LOD, color = epoch_grp)) + 
		geom_step() +
		geom_hline(data = ~.x %>% distinct(epoch_grp, chr, lod_thresh),
				   aes(yintercept = lod_thresh), linetype = 'dashed') + 
		facet_grid(epoch_grp~chr, scales = 'free_x', switch = 'x') + 
		scale_x_continuous(breaks = scales::breaks_pretty(n = 2), 
						   guide = guide_axis(angle = 60)) +
		scale_color_brewer(palette = 'Dark2') + 
		# coord_cartesian(ylim = c(0, 4)) +
		theme_half_open() +
		theme(
			panel.spacing.x = unit(0, 'pt'),
			axis.title.x = element_blank(),
			strip.placement = 'outside',
			strip.text.y = element_text(angle = 0),
			panel.border = element_rect(color = 'gray70'),
			axis.text.x = element_blank(),
			axis.ticks.x = element_blank(),
			legend.position = 'none'
		)
	p
	ggsave(path('pdfs/', 'Suppl_FigEpochScan.pdf'), p, w = 16, h = 8)
```

# Plot founder gt at locus vs. expansion propensity 
```{r fig.width = 6, fig.height = 4, fig.align = 'center', cache = TRUE, eval = TRUE}
# founder genotypes at max LOD marker for %expanded 
fou_gt_at_peak = maxmarg(qtl_data$snp_probs[,'13'], minprob = 0.5) %>% 
		.[[1]] %>%
		as.data.frame %>%
		rownames_to_column(var = 'strain') %>%
		pivot_longer(cols = !strain, names_to = 'marker', values_to = 'fou_gt') %>%
		semi_join(final_qtl_res %>% 
				filter(metric == '% expanded') %>%
				unite('marker', c('chr', 'pos', 'end')) %>%
				slice_max(LOD, n = 1), by = 'marker') %>%
		mutate(fou_gt = recode(fou_gt, `1` = 'B', `2` = 'D'),
			   fou_gt = replace_na(fou_gt, 'miss'))

# Get perc expanded and hap labels for each strain
perc_exp_byhap = all_phenos %>% 
  filter(min_pts_per_phe == 10, max_denovo_strains_per_loc == 10, motif_len == FINALML) %>%
		    filter(metric == '% expanded') %>%
		    filter(strain %in% strains) %>%
		left_join(strain_info %>% 
				  mutate(off_epoch = recode(off_epoch, epoch_1b = 'epoch_1a', epoch_1c = 'epoch_1a')) %>%
				  select(strain = bxd_id, off_epoch), by = 'strain') %>%
		rename(perc_expand = pheno) %>%
		left_join(fou_gt_at_peak %>% select(!marker), by = 'strain') %>%
		arrange(desc(perc_expand)) %>%
		select(strain, perc_expand, off_epoch, fou_gt, n)

founder_colors = c(
		'B' = 'red', 
		'D' = 'blue'
		)
p = perc_exp_byhap %>%
  ggplot(aes(fou_gt, perc_expand, color = fou_gt)) +
  geom_quasirandom(dodge.width = 0.1, width=0.2, shape=21, groupOnX = TRUE) +
  geom_boxplot(aes(group = fou_gt), 
					 fill = 'white', alpha=0, 
					 width = 0.2, outlier.shape = NA,
					 position = position_dodge(width = 0.3)) +
  scale_color_manual(
			values = founder_colors,
			guide = guide_legend(title = NULL, nrow = 1)
		) + 
  theme_half_open() +
  ylab("Expansion propensity")
p
ggsave(path('pdfs/', 'Fig2_SignalByFounder.pdf'), p, w = 6, h = 8)

```

# Plot founder gt at locus vs. expansion propensity 
```{r fig.width = 6, fig.height = 4, fig.align = 'center', cache = TRUE, eval = TRUE}
# founder genotypes at max LOD marker for %expanded 
fou_gt_at_peak = maxmarg(qtl_data$snp_probs[,'13'], minprob = 0.5) %>% 
		.[[1]] %>%
		as.data.frame %>%
		rownames_to_column(var = 'strain') %>%
		pivot_longer(cols = !strain, names_to = 'marker', values_to = 'fou_gt') %>%
		semi_join(final_qtl_res %>% 
				filter(metric == '% expanded') %>%
				unite('marker', c('chr', 'pos', 'end')) %>%
				slice_max(LOD, n = 1), by = 'marker') %>%
		mutate(fou_gt = recode(fou_gt, `1` = 'B', `2` = 'D'),
			   fou_gt = replace_na(fou_gt, 'miss'))

# Get perc expanded and hap labels for each strain
perc_exp_byhap = all_phenos %>% 
  filter(min_pts_per_phe == 10, max_denovo_strains_per_loc == 10, motif_len == FINALML) %>%
		    filter(metric == '% expanded') %>%
		    filter(strain %in% strains) %>%
		left_join(strain_info %>% 
				  mutate(off_epoch = recode(off_epoch, epoch_1b = 'epoch_1a', epoch_1c = 'epoch_1a')) %>%
				  select(strain = bxd_id, off_epoch), by = 'strain') %>%
		rename(perc_expand = pheno) %>%
		left_join(fou_gt_at_peak %>% select(!marker), by = 'strain') %>%
		arrange(desc(perc_expand)) %>%
		select(strain, perc_expand, off_epoch, fou_gt, n)

founder_colors = c(
		'B' = 'red', 
		'D' = 'blue'
		)
p = perc_exp_byhap %>%
  ggplot(aes(fou_gt, perc_expand, color = fou_gt)) +
  geom_quasirandom(dodge.width = 0.1, width=0.2, shape=21, groupOnX = TRUE) +
  geom_boxplot(aes(group = fou_gt), 
					 fill = 'white', alpha=0, 
					 width = 0.2, outlier.shape = NA,
					 position = position_dodge(width = 0.3)) +
  scale_color_manual(
			values = founder_colors,
			guide = guide_legend(title = NULL, nrow = 1)
		) + 
  theme_half_open() +
  ylab("Expansion propensity")
p
ggsave(path('pdfs/', 'Fig2_SignalByFounder.pdf'), p, w = 6, h = 8)

```
# Plot founder gt at locus vs. expansion propensity by epoch
```{r fig.width = 6, fig.height = 4, fig.align = 'center', cache = TRUE, eval = TRUE}
p = perc_exp_byhap %>%
  mutate(off_epoch = str_replace(off_epoch, 'epoch_', '')) %>%
  ggplot(aes(off_epoch, perc_expand, color = fou_gt)) +
  geom_quasirandom(dodge.width = 0.9, width=0.3, shape=21, groupOnX = TRUE) +
  geom_boxplot(
					 fill = 'white', alpha=0, lwd=0.5,
					 width = 0.3, outlier.shape = NA,
					 position = position_dodge(width = 0.9)) +
  scale_color_manual(
			values = founder_colors,
			guide = guide_legend(title = NULL, nrow = 1)
		) + 
  theme_half_open() +
  theme(
			  # legend.position = c(1, 1),
			  # legend.justification = c(1, 1),
			  # legend.text = element_text(size = 8)
			  legend.position = 'none'
		) +
  ylab("Expansion propensity")
p
ggsave(path('pdfs/', 'Fig2_SignalByFounderByEpoch.pdf'), p, w = 16, h = 8)

```

# Plot founder gt at locus vs. contraction size
```{r fig.width = 6, fig.height = 4, fig.align = 'center', cache = TRUE, eval = TRUE}
# founder genotypes at max LOD marker for %expanded 
fou_gt_at_peak = maxmarg(qtl_data$snp_probs[,'13'], minprob = 0.5) %>% 
		.[[1]] %>%
		as.data.frame %>%
		rownames_to_column(var = 'strain') %>%
		pivot_longer(cols = !strain, names_to = 'marker', values_to = 'fou_gt') %>%
		semi_join(final_qtl_res %>% 
				filter(metric == "delta (RU) contr") %>%
				unite('marker', c('chr', 'pos', 'end')) %>%
				slice_max(LOD, n = 1), by = 'marker') %>%
		mutate(fou_gt = recode(fou_gt, `1` = 'B', `2` = 'D'),
			   fou_gt = replace_na(fou_gt, 'miss'))

# Get perc expanded and hap labels for each strain
contradelta_exp_byhap = all_phenos %>% 
  filter(min_pts_per_phe == 10, max_denovo_strains_per_loc == 10, motif_len == FINALML) %>%
		    filter(metric == "delta (RU) contr") %>%
		    filter(strain %in% strains) %>%
		left_join(strain_info %>% 
				  mutate(off_epoch = recode(off_epoch, epoch_1b = 'epoch_1a', epoch_1c = 'epoch_1a')) %>%
				  select(strain = bxd_id, off_epoch), by = 'strain') %>%
		rename(cd = pheno) %>%
		left_join(fou_gt_at_peak %>% select(!marker), by = 'strain') %>%
		arrange(desc(cd)) %>%
		select(strain, cd, off_epoch, fou_gt, n)

founder_colors = c(
		'B' = 'red', 
		'D' = 'blue'
		)
p = contradelta_exp_byhap %>%
  ggplot(aes(fou_gt, cd, color = fou_gt)) +
  geom_quasirandom(dodge.width = 0.1, width=0.2, shape=21, groupOnX = TRUE) +
  geom_boxplot(aes(group = fou_gt), 
					 fill = 'white', alpha=0, 
					 width = 0.2, outlier.shape = NA,
					 position = position_dodge(width = 0.3)) +
  scale_color_manual(
			values = founder_colors,
			guide = guide_legend(title = NULL, nrow = 1)
		) + 
  theme_half_open() +
  ylab("contract_delta_ru")
p
```

# Determine the QTL confidence interval - chr13 % expanded

#. Confidence interval for the QTL determined by the 1.5 LOD drop method - all

```{r echo=TRUE, fig.width = 8, fig.height = 4, fig.align = 'center', cache = TRUE, eval = TRUE}
	# get the confidence interval
  qtl_res_raw = map(final_phenos_qtl, 'qtl_res')
	qtl_ci = map_df(qtl_res_raw, 
					~.x %>% lod_int(map = qtl_data$snp_pmap, chr = '13', threshold = 5, drop = 1.5) %>% as_tibble, 
					.id = 'metric'
					) %>% filter(metric == '% expanded')
	qtl_ci

	p = final_qtl_res %>%
		mutate(chr = str_replace(chr, 'chr', '')) %>%
		mutate(chr = fct_relevel(chr, str_sort(unique(chr), numeric = TRUE))) %>%
		filter(chr == '13') %>%
		mutate(across(c(pos, end), ~.x/1e6)) %>%
		ggplot(aes(pos, LOD), by=metric) + 
		geom_rect(aes(xmin = qtl_ci$ci_lo, xmax = qtl_ci$ci_hi, ymin = -Inf, ymax = Inf), fill = 'gray90') + 
		geom_step() +
		geom_hline(data = ~.x %>% distinct(metric, chr, lod_thresh),
				   aes(yintercept = lod_thresh), linetype = 'dashed') +
		facet_grid(metric~chr, scales = 'free_x') + 
		scale_x_continuous(breaks = scales::breaks_width(width = 5), 
						   guide = guide_axis(angle = 60)) +
		theme_half_open() +
		theme(
			panel.spacing.x = unit(0, 'pt'),
			axis.title.x = element_blank(),
			strip.placement = 'outside',
			strip.text.y = element_text(angle = 0),
			panel.border = element_rect(color = 'gray70'),
			# axis.text.x = element_blank(),
			# axis.ticks.x = element_blank(),
			legend.position = 'top'
		)
	p
	
```

#. Confidence interval for the QTL determined by the 1.5 LOD drop method - tetra

```{r echo=TRUE, fig.width = 8, fig.height = 4, fig.align = 'center', cache = TRUE, eval = TRUE}
	# get the confidence interval
  qtl_res_raw = map(final_phenos_qtl_tetra, 'qtl_res')
	qtl_ci_tetra = map_df(qtl_res_raw, 
					~.x %>% lod_int(map = qtl_data$snp_pmap, chr = '13', threshold = 5, drop = 1.5) %>% as_tibble, 
					.id = 'metric'
					) %>% filter(metric == '% expanded')
	qtl_ci_tetra

	p = final_qtl_res_tetra %>%
		mutate(chr = str_replace(chr, 'chr', '')) %>%
		mutate(chr = fct_relevel(chr, str_sort(unique(chr), numeric = TRUE))) %>%
		filter(chr == '13') %>%
		mutate(across(c(pos, end), ~.x/1e6)) %>%
		ggplot(aes(pos, LOD), by=metric) + 
		geom_rect(aes(xmin = qtl_ci_tetra$ci_lo, xmax = qtl_ci_tetra$ci_hi, ymin = -Inf, ymax = Inf), fill = 'gray90') + 
		geom_step() +
		geom_hline(data = ~.x %>% distinct(metric, chr, lod_thresh),
				   aes(yintercept = lod_thresh), linetype = 'dashed') +
		facet_grid(metric~chr, scales = 'free_x') + 
		scale_x_continuous(breaks = scales::breaks_width(width = 5), 
						   guide = guide_axis(angle = 60)) +
		theme_half_open() +
		theme(
			panel.spacing.x = unit(0, 'pt'),
			axis.title.x = element_blank(),
			strip.placement = 'outside',
			strip.text.y = element_text(angle = 0),
			panel.border = element_rect(color = 'gray70'),
			# axis.text.x = element_blank(),
			# axis.ticks.x = element_blank(),
			legend.position = 'top'
		)
	p
	
```

# Plot traces by epoch for chr13 % expanded locus - TODO - maybe in python

#. might do this in Python
#. Can include pre-specified list of genes
#. get byepoch from perc_expand_by_epoch_final
#. get overall from final_qtl_res, final_qtl_res_tetra
#. plot region corresponding to the 1.5-LOD (~75-97 Mb)
#. show separate confidence interval for tetra-only