---
title: "Probe level eQTL analysis"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
      df_print: paged
      code_folding: hide
      toc: true
      toc_float: true
      number_sections: true
---

<!-- Set with output width -->
<style type="text/css">
div.main-container {
    max-width: 2000px;
}
.col-md-3 {
    width: 15%;
}
h1 {font-size: 2em}
h2 {font-size: 1.5em}
h3 {font-size: 1.17em}
h4 {font-size: 1.12em}
h5 {font-size: 0.83em}
h6 {font-size: 0.75em}
</style>

```{r setup, include=FALSE}
	# options
	knitr::opts_chunk$set(echo = TRUE)
	options(stringsAsFactors = FALSE)

	# libraries
	library(tidyverse)
	library(cowplot)
	library(fs)
	library(ggrepel)
	library(flextable)
	library(jsonlite)
	library(ggbeeswarm)
	library(ggh4x)

```

# Plotting functions

```{r echo=TRUE, fig.width = 6, fig.height = 4, fig.align = 'center', cache = TRUE, eval = TRUE}
	# plot trascript tracks
	tx_track = function(data, xlims = NULL, ylabs = FALSE) {
		p = data %>%
			arrange(tx_type) %>% 
			mutate(tx_id = fct_inorder(tx_id)) %>%
			ggplot() +
			geom_tile(data = ~.x %>% distinct(tx_id, exon_pos, exon_end),
				  aes(width = (exon_end - exon_pos), y = tx_id, x = (exon_pos + exon_end)/2), height = 0.9, fill = NA, color = 'black') +
			geom_tile(data = ~.x %>% distinct(tx_id, cds_pos, cds_end) %>% filter(!is.na(cds_pos)),
				  aes(width = (cds_end - cds_pos), y = tx_id, x = (cds_pos + cds_end)/2), height = 0.9) +
			geom_tile(data = ~.x %>% distinct(tx_id, tx_pos, tx_end, tx_type),
				  aes(width = (tx_end - tx_pos), y = tx_id, x = (tx_pos + tx_end)/2, fill = tx_type), height = 0.1) +
			# geom_point(data = ~.x %>% distinct(tx_id, tss),
			# 	   aes(x = tss, y = tx_id)) + 
			scale_fill_brewer(palette = 'Dark2') + 
			theme_half_open() +
			labs(x = 'pos')
		if (!is.null(xlims)) p = p + coord_cartesian(xlim = xlims) 
		if (ylabs) {
			p = p + theme(legend.position = 'none',
				  axis.title.y = element_blank())
		} else {
			p = p + theme(legend.position = 'none',
				  axis.text.y = element_blank(),
				  axis.title.y = element_blank())
		}
		return(p)
	}
	
	# plotting function
	plt_by_X = function(data, color_by = 'GN', ycol) {
		p_list = map(dna_repair_genes, function(.x) {
			xlims = data %>% filter(gene_name == .x) %>% pull(probe_mid) %>% range
			xlims = xlims/1e6
			p1 = tx_info %>% 
				filter(gene_name == .x) %>%
				mutate(across(matches('_(pos|end)'), ~.x/1e6)) %>%
				tx_track(xlims = xlims) +
				facet_wrap(~gene_name)
			p2 = data %>%
				mutate(probe_mid = probe_mid/1e6) %>%
				filter(gene_name == .x) %>%
				ggplot(aes_string('probe_mid', ycol)) + 
				geom_point(aes_string(color = color_by)) + 
				scale_color_viridis_d(na.value = 'gray40', guide = guide_legend(nrow = 5)) + 
				coord_cartesian(xlim = xlims) +
				theme_half_open() + 
				theme(legend.position = 'none')
			plot_grid(p1, p2, ncol = 1, axis = 'lr', align = 'v')
		})
		lgnd = ( data %>% 
			ggplot(aes_string('probe_mid', ycol)) + 
			geom_point(aes_string(color = color_by)) + 
			scale_color_viridis_d(guide = guide_legend(ncol = 4)) + 
			theme_half_open() + 
			theme(legend.position = 'right', 
				  legend.key.size = unit(0.1, 'cm'), 
				  legend.text = element_text(size = 6)) 
		) %>% get_legend()
		p = plot_grid(plot_grid(plotlist = p_list, nrow = 1), lgnd, nrow = 1, rel_widths = c(1, 0.15))
		p
	}
	
		
	# plotting function
	plt_by_nvars = function(data, ycol) {
		p_list = map(dna_repair_genes, function(.x) {
			xlims = data %>% filter(gene_name == .x) %>% pull(probe_mid) %>% range
			xlims = xlims/1e6
			p1 = tx_info %>% 
				filter(gene_name == .x) %>%
				mutate(across(matches('_(pos|end)'), ~.x/1e6)) %>%
				tx_track(xlims = xlims, ylabs = TRUE) +
				facet_wrap(~gene_name)
			
			p2 = data %>%
				mutate(n_var_per_probe = as.factor(n_var_per_probe)) %>%
				mutate(probe_mid = probe_mid/1e6) %>%
				filter(gene_name == .x) %>%
				ggplot(aes_string('probe_mid', ycol)) + 
				geom_point(data = ~.x %>% filter(n_var_per_probe == 0), color = 'gray70') + 
				geom_point(data = ~.x %>% filter(n_var_per_probe != 0), aes(color = n_var_per_probe)) + 
				scale_color_viridis_d(option = 'plasma', guide = guide_legend(ncol = 3)) + 
				coord_cartesian(xlim = xlims) +
				theme_half_open() + 
				theme(legend.position = c(0.05, 1),
						legend.title = element_text(size = 8),
						legend.justification = c(0, 1),
						legend.text = element_text(size = 8),
						legend.key.size = unit(0.1, 'cm'))
			plot_grid(p1, p2, ncol = 1, axis = 'lr', align = 'v')
		})
		p = plot_grid(plotlist = p_list, nrow = 1)
		p
	}
```

# Load eQTL and gene data

```{r echo=TRUE, fig.width = 6, fig.height = 4, fig.align = 'center', cache = TRUE, eval = TRUE}
	# load full eqtl results and probe information
	eqtl_data  = readRDS('../data/gene_expr/qtl_agg/gene_expr_db.rds')
	probe_info = eqtl_data$probes # %>% mutate(across(matches('_(pos|end)'), ~.x*1e6))
	traces     = eqtl_data$signal

	# transcript info
	tx_info = readRDS('../data/analysis_cache/annot/vep_data.rds')$tx_info
	tx_info = tx_info %>% mutate(across(matches('_(pos|end)'), ~.x*1e6 %>% as.integer))

	# reload list of genes from EMBL
	embl_genes = readRDS('../data/analysis_cache/embl_genes.rds')
	
	# get protein coding genes
	prot_genes = embl_genes %>% filter(gene_type == 'protein_coding') %>% pull(gene_name)

	# select genes of interest
	dna_repair_genes = c('Msh3', 'Atg10', 'Xrcc4', 'Ssbp2')

	# load eqtl data where snps are used as covariates for probes with snps
	traces_snps_covar = readRDS('../data/gene_expr/qtl_agg_snps_covar/gene_expr_db.rds')$signal

	# join LOD values from the snps covar onto the traces
	traces = traces %>%
		left_join(traces_snps_covar %>% rename(LOD_snps_covar = LOD, thresh_snps_covar = thresh),
				  by = c('GN', 'marker', 'probe', 'gene_name'))

	# check
	if (0) {
		traces %>% 
			left_join(probe_info, by = 'probe') %>%
			count(n_var_per_probe == 0, is.na(LOD_snps_covar))
		# `n_var_per_probe == 0` `is.na(LOD_snps_covar)`       n
		# <lgl>                  <lgl>                     <int>
		# FALSE                  FALSE                    801900
		# TRUE                   TRUE                    4173768
	}

	# replace missing LOD_snps_covar values with LOD values for probes that have no snps
	traces = traces %>% 
		mutate(LOD_snps_covar    = if_else(is.na(LOD_snps_covar), LOD, LOD_snps_covar),
			   thresh_snps_covar = if_else(is.na(thresh_snps_covar), thresh, thresh_snps_covar))

	# load analyzed eQTL data (representative datasets, eqtl_dset_probe combinations etc.
	# NOTE: this is mostly for main figures
	proc_eqtl_data = readRDS('../data/analysis_cache/eqtl_data.rds')
	eqtl_dsets_genes = proc_eqtl_data$eqtl_dsets_genes
	sel_dsets = proc_eqtl_data$sel_dsets

	# load special detail eqtl mapping results for all loci with QTL region for Msh3 on
	msh3_eqtl_detailed = readRDS('../data/analysis_cache/eqtl_detailed.rds')
```

# Keep top eQTL value per trace

#. Keep top LOD value per probe
#. Don't need LOD values for every marker

```{r echo=TRUE, fig.width = 6, fig.height = 4, fig.align = 'center', cache = TRUE , eval = TRUE}
	# this is best point per GN/probe/gene_name
	# calling it probe_max to distinguish from the "best_point" variant I use to refers to max per GN/gene
	probe_max = traces %>%
		filter(gene_name %in% prot_genes) %>%
		left_join(probe_info, by = 'probe') %>%
		arrange(desc(LOD)) %>%
		distinct(GN, probe, gene_name, .keep_all = TRUE)
	probe_max_snps_covar = traces %>%
		filter(gene_name %in% prot_genes) %>%
		left_join(probe_info, by = 'probe') %>%
		arrange(desc(LOD_snps_covar)) %>%
		distinct(GN, probe, gene_name, .keep_all = TRUE)
```

# Probes are grouped by type

#. RNAseq datasets
	- VCU_BXD_PFC_Et_vs_Sal RNAseq datasets:  GN900, GN899, GN898, GN884, GN885, GN883
	- RNAseq on ABI SOLiD from UTSCH_Mouse_BXD_Whole_Brain: GN394, GN164, GN590, GN589
#. There are also a number of proteomics datasets which have peptide fragments instead of probes
#. RNAseq and proteomics datasets should be treated differently from mRNA probe arrays

```{r echo=TRUE, fig.width = 6, fig.height = 4, fig.align = 'center', cache = TRUE, eval = TRUE}
	if (0) {
		# probes that look like gene ids
		probe_max %>%
			filter(str_detect(probe, 'ENSMUSG')) %>%
			pull(GN) %>% unique
		# [1] GN900 GN899 GN898 GN884 GN885 GN883
		# These are from the VCU_BXD_PFC_Et_vs_Sal dataset that is actually RNAseq
		# this should be a priveleged dataset because it wouldn't be susceptible to probe artifacts
		# Also this dataset doesn't have Msh3, but has other Msh proteins

		# probes that look like transcript ids 
		probe_max %>%
			filter(str_detect(probe, 'NM_')) %>%
			pull(GN) %>% unique
		# [1] GN394 GN164 GN590 GN589
		# These are from 
		# this was done on ABI SOLiD
		# similar to RNA seq
	}
	
	# make a distinct list of probes
	probes = probe_max %>% distinct(gene_name, across(matches('probe')), BlatSeqLen, alig_data)

	# remove alig_data from probe_max and probe_max_snps_covar
	probe_max$alig_data = NULL
	probe_max_snps_covar$alig_data = NULL
	
	# check counts
	probes %>% count(probe_type)
	
	# NOTE: that probe_len_blat_bp is same as sum of individual alignment lengths
	if (0) {
		# calculate total length of aligned BLAT fragments
		probes %>%
			filter(is_probeset) %>%
			select(probe, probe_len_blat_bp, alig_data) %>%
			unnest(alig_data) %>%
			mutate(len = alig_end - alig_pos) %>%
			group_by(probe, probe_len_blat_bp) %>%
			summarise(tot_len = sum(len), .groups = 'drop') %>%
			count(tot_len == probe_len_blat_bp)

		# check on position
		probes %>%
			filter(is_probeset) %>%
			select(probe, probe_pos, probe_end, alig_data) %>%
			unnest(alig_data) %>%
			mutate(mid = (alig_end + alig_pos)/2) %>%
			group_by(probe, probe_pos, probe_end) %>%
			summarise(mid = mean(mid), .groups = 'drop') %>%
			mutate(probe_mid = (probe_end + probe_pos)/2) %>%
			mutate(diff = probe_mid - mid) %>%
			skimr::skim(diff)
	}

	# calculate the probe_mid based on alignments
	probe_mids = probes %>%
		filter(is_probeset) %>%
		select(probe, alig_data) %>%
		unnest(alig_data) %>%
		mutate(mid = (alig_end + alig_pos)/2) %>%
		group_by(probe) %>%
		summarise(probe_mid = mean(mid), .groups = 'drop')
	probes = probes %>% left_join(probe_mids, by = 'probe') %>% select(!alig_data)

	# join the probe mids to probe_max and probe_max_snps_covar
	probe_max = probe_max %>%
		left_join(probe_mids, by = 'probe') %>%
		mutate(probe_mid = if_else(is.na(probe_mid), (probe_pos+probe_end)/2, probe_mid))
	probe_max_snps_covar = probe_max_snps_covar %>%
		left_join(probe_mids, by = 'probe') %>%
		mutate(probe_mid = if_else(is.na(probe_mid), (probe_pos+probe_end)/2, probe_mid))
```


# Final check on probe counts

```{r echo=TRUE, fig.width = 6, fig.height = 4, fig.align = 'center', cache = TRUE, eval = TRUE}
	probes %>% count(probe_type, is_probeset, n_probes > 1, name = 'n_probe')
```


# Variants per probe (corrected LOD)

#. Array probes only
#. LOD corrected by using number of snps per variant as covariate

```{r echo=TRUE, fig.width = 17, fig.height = 6, fig.align = 'center', cache = TRUE, eval = TRUE, warning = FALSE}
	p = probe_max_snps_covar %>%
		filter(probe_type == 'probe_array') %>%
		plt_by_nvars(ycol = 'LOD_snps_covar')
	p
```

```{r echo=TRUE, fig.width = 17, fig.height = 6, fig.align = 'center', cache = TRUE, eval = TRUE, warning = FALSE}
  xlims = c(92.211872, 92.355000)
  dna_repair_genes = c('Msh3')
	p = probe_max_snps_covar %>%
		filter(probe_type == 'probe_array') %>%
		plt_by_nvars(ycol = 'LOD_snps_covar')
	p
	ggsave('../pdfs/SuppFig18_probe-levelanalysis.pdf', p, w = 12, h = 6)

```

# Variants per probe (corrected LOD & passing threshold)

```{r echo=TRUE, fig.width = 17, fig.height = 6, fig.align = 'center', cache = TRUE, eval = TRUE, warning = FALSE}
	dna_repair_genes = c('Msh3', 'Atg10', 'Xrcc4', 'Ssbp2')
	p = probe_max_snps_covar %>%
		filter(probe_type == 'probe_array') %>%
		filter(LOD_snps_covar >= thresh_snps_covar) %>%
		plt_by_nvars(ycol = 'LOD_snps_covar')
	p
```


# MSH3 by probe eQTL plots

## Full

```{r echo=TRUE, fig.width = 9, fig.height = 5.5, fig.align = 'center', cache = TRUE, eval = TRUE}
	xlims = c(92.211872, 92.355000)
	# transcripts
	msh3_tx = tx_track(
		tx_info %>% filter(gene_name == 'Msh3') %>% mutate(across(matches('_(pos|end)'), ~.x/1e6)),
		xlims = xlims, ylabs = TRUE)
	
	# manhattan style
	msh3_lod = probe_max_snps_covar %>%
		filter(gene_name == 'Msh3') %>%
		mutate(across(matches('_(pos|end|mid)'), ~.x/1e6)) %>%
		group_by(GN, gene_name) %>%
		mutate(has_eqtl = any(LOD_snps_covar >= thresh_snps_covar)) %>%
		ungroup %>%
		filter(has_eqtl) %>%
		ggplot(aes(probe_mid, LOD_snps_covar)) +
		geom_point(color = 'gray50') +
		# geom_point(aes(color = GN)) +
		scale_color_viridis_d(na.value = 'gray40') +
		theme_half_open() +
		theme(legend.position = 'left') +
		coord_cartesian(xlim = xlims)
	# visualize location of probes
	msh3_probes = probe_max_snps_covar %>%
		filter(gene_name == 'Msh3') %>%
		filter(probe_type == 'probe_array') %>%
		group_by(probe_chr, probe_mid, probe_pos, probe_end, probe_len_bp) %>%
		# take max LOD per probe across GNs
		summarise(maxLOD = max(LOD, na.rm = TRUE), .groups = 'drop') %>%
		mutate(y = 0.1*IRanges::disjointBins(IRanges::IRanges(probe_pos, probe_end))) %>%
		mutate(across(c(probe_len_bp, matches('_(pos|end|mid)')), ~.x/1e6)) %>%
		# mutate(probe_mid = (probe_end + probe_pos)/2) %>% 
		ggplot() + 
		geom_tile(aes(x = probe_mid, width = probe_len_bp, y = y, fill = maxLOD), height = 0.05) +  
		scale_fill_viridis_c() + 
		coord_cartesian(xlim = xlims) +
		theme_half_open() + 
		theme(axis.text.y = element_blank(),
			  axis.title.y = element_blank(),
			  axis.ticks.y = element_blank(),
			  legend.position = 'left'
		)
	p = plot_grid(msh3_tx, msh3_lod, ncol = 1, align = 'v', axis = 'lr', rel_heights = c(1.2, 1))
	p
```

## Zoom w/ probe positions

```{r echo=TRUE, fig.width = 10, fig.height = 6.5, fig.align = 'center', cache = TRUE, eval = TRUE}
	xlims = c(92.34, 92.355)
	p1 = msh3_tx + coord_cartesian(xlim = xlims)
	p2 = msh3_lod + coord_cartesian(xlim = xlims) 
	p3 = msh3_probes + coord_cartesian(xlim = xlims)
	p = plot_grid(p1, p2, p3, ncol = 1, align = 'v', axis = 'lr', rel_heights = c(1.5, 1, 1))
	p
	# ggsave('test.pdf', p, w = 10, h = 6.5)
```


# Supplementary figure 14

```{r echo=TRUE, fig.width = 14, fig.height = 9, fig.align = 'center', cache = TRUE, eval = TRUE}
	xlims = c(92.211872, 92.355000)
	xlims_zoom = c(92.345, 92.355)

	# transcripts
	p1 = tx_track(
		tx_info %>% 
			filter(gene_name == 'Msh3') %>% 
			# filter(tx_id == 'ENSMUST00000185852.6') %>%
			filter(tx_id %in% c('ENSMUST00000185852.6', 'ENSMUST00000190393.6')) %>%
			mutate(across(matches('_(pos|end)'), ~.x/1e6)),
		xlims = xlims, ylabs = TRUE) + 
		theme(axis.text.y = element_blank())

	# eqtls by probe
	p2 = probe_max_snps_covar %>%
		filter(gene_name == 'Msh3') %>%
		# NOTE: keep only datasets where gene is significant, but not top probe
		semi_join(eqtl_dsets_genes, by = c('GN', 'gene_name')) %>%
		mutate(across(matches('_(pos|end|mid)'), ~.x/1e6)) %>%
		mutate(probe = as.character(probe)) %>%
		ggplot(aes(probe_mid, LOD_snps_covar)) + 
		# geom_point(color = 'gray50') + 
		geom_point(aes(color = GN)) + 
		geom_text_repel(aes(label = if_else(probe %in% c('1446511_at', '17294887', '4933342', 'A_55_P1974477'), probe, '')),
						box.padding = 0.6,
						min.segment.length = 0) +
		scale_color_viridis_d(na.value = 'gray40') + 
		theme_half_open() + 
		theme(
			  # legend.position = 'left'
			  legend.position = c(0.025, 1),
			  legend.justification = c(0, 1),
			  legend.background = element_rect(fill = 'white', color = 'black'),
			  legend.key.size = unit(1, 'pt'),
			  legend.text = element_text(size = 8),
			  legend.title = element_text(size = 8),
			  legend.margin = margin(t = 3, b = 3, l = 3, r = 3, unit = 'pt')
		) +
		coord_cartesian(xlim = xlims) +
		labs(y = 'LOD')

	## with individual snps
	p3 = msh3_eqtl_detailed %>%
		filter(!is.na(p.value)) %>%
		separate('loc_id', c('chr', 'pos', 'end', 'sv_type'), convert = TRUE, fill = 'right') %>%
		mutate(across(c(pos, end), ~.x/1e6)) %>%
		mutate(LOD = -log10(p.value)) %>%
		ggplot(aes(pos, LOD)) + 
		geom_point(aes(color = probe)) + 
		# geom_point(color = 'gray50') + 
		scale_color_viridis_d(na.value = 'gray40', guide = guide_legend(nrow = 1, title.position = 'left')) + 
		theme_half_open() + 
		theme(
			  # legend.position = 'left'
			  legend.position = c(0.025, 1),
			  legend.justification = c(0, 1),
			  legend.background = element_rect(fill = 'white', color = 'black'),
			  legend.key.size = unit(1, 'pt'),
			  legend.text = element_text(size = 8),
			  legend.title = element_text(size = 8),
			  legend.margin = margin(t = 3, b = 3, l = 3, r = 3, unit = 'pt')
		) +
		coord_cartesian(xlim = xlims) +
		labs(y = '-log10(p.value)')

	hlite = geom_rect(xmin = xlims_zoom[1], xmax = xlims_zoom[2], ymin = 0, ymax = Inf, fill = 'gray70', alpha = 0.01)
	p_a = plot_grid(ggdraw() + draw_text('By probe' , x = 0.1, y = 0.5, fontface = 'bold'), 
					p1 + hlite, 
					p2 + hlite, 
					ncol = 1, align = 'v', axis = 'lr', rel_heights = c(0.1, 0.3, 1, 1))
	p_b = plot_grid(ggdraw() + draw_text('By marker', x = 0.1, y = 0.5, fontface = 'bold'), 
					p1 + hlite, 
					p3 + hlite, 
					ncol = 1, align = 'v', axis = 'lr', rel_heights = c(0.1, 0.3, 1, 1))
	p_c = plot_grid(ggdraw() + draw_text('By probe (zoom)' , x = 0.1, y = 0.5, fontface = 'bold'), 
					p1 + coord_cartesian(xlim = xlims_zoom), 
					p2 + coord_cartesian(xlim = xlims_zoom),
					ncol = 1, align = 'v', axis = 'lr', rel_heights = c(0.1, 0.3, 1, 1))
	p_d = plot_grid(ggdraw() + draw_text('By marker (zoom)', x = 0.1, y = 0.5, fontface = 'bold'), 
					p1 + coord_cartesian(xlim = xlims_zoom), 
					p3 + coord_cartesian(xlim = xlims_zoom),
					ncol = 1, align = 'v', axis = 'lr', rel_heights = c(0.1, 0.3, 1, 1))
	p = plot_grid(p_a, p_b, p_c, p_d, nrow = 2)
	p

	w = 20; h = 9

	# save figure
	ggsave('../pdfs/SuppFig22_Detailedprobe-levelanalysisofmsh3.pdf', p, w = w, h = 9)
```
